---
title: "Beta Diversity and Differential Abundance Tests for the North Pond FLOCs experiments"
author: "Tim D'Angelo"
date: "4/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(phyloseq)
library(decontam)
library(ALDEx2)
library(zCompositions)
library(vegan)
```


```{r}
ps <- readRDS(file="NP_FLOCSv138_ps") # file seqproc file, updated with Silva v138 taxonomy
meta <- read.csv("flocs-meta.csv", row.names = 1)
pm <- sample_data(meta)
ps_m <- merge_phyloseq(ps, pm)
```


```{r}
npfa_rel <- transform_sample_counts(ps_m, function(x) x/sum(x)) # convert to relative abundance for Decontam 
```

```{r}
sample_data(npfa_rel)$is.neg <- sample_data(npfa_rel)$decontam == "NTC"
contamdf.prev025 <- isContaminant(npfa_rel, method="prevalence", neg="is.neg", threshold=0.25)
table(contamdf.prev025$contaminant)
```

ASVs flagged as contaminants by Decontam. Manualy removed ASV_11, annotated as Ralstonia. It is not flagged by Decontam but is suspicious because it is in all NTCs
```{r}
npfa_36cont <- c("ASV_162", "ASV_165", "ASV_328", "ASV_344", "ASV_400", "ASV_411", "ASV_415", "ASV_426", "ASV_450", "ASV_512", "ASV_524", "ASV_689", "ASV_690", "ASV_692", "ASV_818", "ASV_904", "ASV_950", "ASV_986", "ASV_1010", "ASV_1018", "ASV_1035", "ASV_1156", "ASV_1312", "ASV_1365", "ASV_1368", "ASV_1561", "ASV_1654", "ASV_1760", "ASV_1825", "ASV_2142", "ASV_2143", "ASV_2505", "ASV_2507", "ASV_2690", "ASV_2854", "ASV_3314", "ASV_11")
```


```{r}
pop_taxa = function(physeq, badTaxa){
  allTaxa = taxa_names(physeq)
  myTaxa <- allTaxa[!(allTaxa %in% badTaxa)]
  return(prune_taxa(myTaxa, physeq))
}
```

```{r}
npfa_nocont = pop_taxa(ps_m, npfa_36cont)
npfa_nocont
```

```{r}
npfa_nocontSO <- subset_samples(npfa_nocont, decontam =="sample") # remove NTC samples
npfa_nocontSO
```


Remove ASVs with count = 0. These are ASVs only present in the electrochemistry dataset. 
```{r}
wh0 = genefilter_sample(npfa_nocontSO, filterfun_sample(function(x) x > 0))
npfa_nocontPR = prune_taxa(wh0, npfa_nocontSO)
npfa_nocontPR
```



DADA2 does not produce singletons, the minimum count of an ASV is 2. Here ASVs with counts of 2 are removed and only ASVs present in 4 or more samples are removed.  This removes a large number of ASVs but a very small fraction of the total reads in the dataset, indicating they are spurious. 
```{r}
wh0 = genefilter_sample(npfa_nocontPR, filterfun_sample(function(x) x > 2), A=0.05*nsamples(npfa_nocontPR))
npfa_pr = prune_taxa(wh0, npfa_nocontPR)
npfa_pr
```


Same thing for downhole samples only.
```{r}
npfa_DH <- subset_samples(npfa_nocontPR, FlocsLoc=="DH")
wh0 = genefilter_sample(npfa_DH, filterfun_sample(function(x) x > 2), A=0.05*nsamples(npfa_DH))
DH_pr = prune_taxa(wh0, npfa_DH)
DH_pr
```


Example code for calculation of Aitchison Distance and Ordination vis Multidimensional Scaleing (Principle Coordinates Analysis)
```{r}
#write.csv(otu_table(npfa_pr), file = "npfa_asv_pr.csv")
#pr_asv <- read.csv("npfa_asv_pr.csv", header = TRUE, row.names = 1)
```

```{r}
d.czm <- cmultRepl(pr_asv,  label=0, method="CZM")
d.clr <- t(apply(d.czm, 1, function(x){log(x) - mean(log(x))}))
a.dist <- dist(d.clr)
```


```{r}
Aitchinson_PCA <- ordinate(npfa_pr, 'MDS', distance=a.dist)
```

Example Code for Differential Tests
```{r}
cork_v_lite.meta <- read.csv(file="cork-corklite-meta.csv", header = TRUE, row.names=1) # metadata delineating U1383C + U1382A from U1382B
write.csv(otu_table(DH_pr), file = "DH_cork_comp.csv") #
DH_cork_asv <- read.csv("DH_cork_comp.csv", header = TRUE, row.names = 1) #
DH_cork_T <- t(DH_cork_asv) # ASV count table in correct format
conds <- as.character(cork_v_lite.meta$CorkType) # define categories
```

Run ALDEX2 Tests. Figures in paper only included ASVs with an effect size greater than 1 and BH corrected p value > 0.05
```{r}
#cork_ALDEX2 <- aldex.clr(DH_cork_T, conds = conds, mc.samples=128)
#cork_effects <- aldex.effect(cork_ALDEX2, conds)
#cork_stat <- aldex.ttest(cork_ALDEX2, conds)
#cork_DAT <- data.frame(cork_effects, cork_stat)
```


